---
title: "Lab9-FA-SEM"
format: html
editor: visual
execute:
  freeze: auto
  echo: true
  message: false
  warning: false
---

# Lab 8 - Factor Analysis and SEM

-   We are going to assess a scale that measures the anxiety that statistics provokes in students

-   The dataset is from Andy Field with the following questions:

    1.  Statistics makes me cry
    2.  My friends will think I'm stupid for not being able to cope with R
    3.  Standard deviations excite me
    4.  I dream that Pearson is attacking me with correlation coefficients
    5.  I don't understand statistics
    6.  I have little experience of computers
    7.  All computers hate me
    8.  I have never been good at mathematics
    9.  My friends are better at statistics than me
    10. Computers are useful only for playing games
    11. I did badly at mathematics at school
    12. People try to tell you that R makes statistics easier to understand but it doesn't
    13. I worry that I will cause irreparable damage because of my incompetence with computers
    14. Computers have minds of their own and deliberately go wrong whenever I use them
    15. Computers are out to get me
    16. I weep openly at the mention of central tendency
    17. I slip into a coma whenever I see an equation
    18. R always crashes when I try to use it
    19. Everybody looks at me when I use R
    20. I can't sleep for thoughts of eigenvectors
    21. I wake up under my duvet thinking that I am trapped under a normal distribution
    22. My friends are better at R than I am
    23. If I'm good at statistics my friends will think I'm a nerd

## Load data

```{r}
devtools::install_github("dr-JT/semoutput")
library(corrplot) # nice correlation plots
library(easystats) # paramaters
library(tidyverse) # tidying visualization
library(psych) # fa func

```

```{r}

data <- read_csv("https://raw.githubusercontent.com/doomlab/learnSEM/master/vignettes/data/lecture_efa.csv")

```

## Exploratory Factor Analysis

1.  Explore the data.

    > Make sure there are no missing data

    ```{r}

    ```

    > Look for outliers and remove them

    ```{r}

    ```

    > Plot the correlation matrix

    ```{r}

    ```

2.  Conduct and interpret the two diagnostic tests to determine if factor analysis is appropriate as an analysis

    ```{r}

    ```

3.  Determine how many factors to extract. Inspect and interpret eigenvalues \> 1, scree plot, parallel plot, and the consensus method.

    ```{r}

    ```

    ```{r}

    ```

    ```{r}

    ```

> Do they all agree? Make a choice on as to how many factors you want to extract

::: callout-note
## ANSWER
:::

1.  Conduct a factor analysis using `psych::fa`(make sure to choose an extraction method and rotation method)

    ```{r}


    ```

2.  Output the factors loadings in a nice table

    ```{r}

    ```

    > Which items appear to load on which factor? Are there any items that seem to load strongly on multiple factors (i.e., items that are cross-loaded)?

    ::: callout-note
    ## ANSWER
    :::

> Use `model_paramaters` and the `threshold` argument to get rid of loadings greater than .32. Does that help with interpretation? If not, try setting the `threshold` to max.

```{r}

```

6.  Come up with names for the factors you extracted.

    ::: callout-note
    ## ANSWER
    :::

## Confirmatory Factor Analysis

7.  Split your data into a testing and training datasets

    ```{r}
    ```

8.  Fit the model you created on the test data using the the `cfa` function. How does the model fit fit? Use `semoutput (https://dr-jt.github.io/semoutput/)`

    ```{r}
    #look at fit indices dicussed in class 

    ```

9.  Fit a competing model (the choice is yours).

    ```{r}

    ```

    > How does this model fit?

    ::: callout-note
    ## ANSWER
    :::

10. Compare both models you fit above. Which one fits the data better?

    ```{r}
    #LRT 
    ```

    ::: callout-note
    ## ANSWER
    :::

11. Write up the results. Be sure to include a figure of the final EFA and a table containing factor loadings, communality, uniqueness, and complexity
