---
title: "Lab 9 - Factor Analysis and CFA"
format: html
editor: visual
execute:
  freeze: auto
  echo: true
  message: false
  warning: false
---

-   We are going to assess a scale that measures the anxiety that statistics provokes in students

-   The dataset is from Andy Field. Each question was a statement followed by a five-point Likert scale: strongly disagree = 1, disagree, neither agree nor disagree, agree and strongly agree (SD, D, N, A and SA respectively).

+------------------------------------------------------------------------------------------------------------------------------------------+
| -   **id**: The student's id                                                                                                             |
|                                                                                                                                          |
| -   **raq_01**: responses (1-5) to the question *Statistics make me cry*                                                                 |
|                                                                                                                                          |
| -   **raq_02**: responses (1-5) to the question *My friends will think I'm stupid for not being able to cope with R*                     |
|                                                                                                                                          |
| -   **raq_03**: responses (1-5) to the question *Standard deviations excite me*                                                          |
|                                                                                                                                          |
| -   **raq_04**: responses (1-5) to the question *I dream that Pearson is attacking me with correlation coefficients*                     |
|                                                                                                                                          |
| -   **raq_05**: responses (1-5) to the question *I don't understand statistics*                                                          |
|                                                                                                                                          |
| -   **raq_06**: responses (1-5) to the question *I have little experience of computers*                                                  |
|                                                                                                                                          |
| -   **raq_07**: responses (1-5) to the question *All computers hate me*                                                                  |
|                                                                                                                                          |
| -   **raq_08**: responses (1-5) to the question *I have never been good at mathematics*                                                  |
|                                                                                                                                          |
| -   **raq_09**: responses (1-5) to the question *My friends are better at statistics than me*                                            |
|                                                                                                                                          |
| -   **raq_10**: responses (1-5) to the question *Computers are useful only for playing games*                                            |
|                                                                                                                                          |
| -   **raq_11**: responses (1-5) to the question *I did badly at mathematics at school*                                                   |
|                                                                                                                                          |
| -   **raq_12**: responses (1-5) to the question *People try to tell you that R makes statistics easier to understand but it doesn't*     |
|                                                                                                                                          |
| -   **raq_13**: responses (1-5) to the question *I worry that I will cause irreparable damage because of my incompetence with computers* |
|                                                                                                                                          |
| -   **raq_14**: responses (1-5) to the question *Computers have minds of their own and deliberately go wrong whenever I use them*        |
|                                                                                                                                          |
| -   **raq_15**: responses (1-5) to the question *Computers are out to get me*                                                            |
|                                                                                                                                          |
| -   **raq_16**: responses (1-5) to the question *I weep openly at the mention of central tendency*                                       |
|                                                                                                                                          |
| -   **raq_17**: responses (1-5) to the question *I slip into a coma whenever I see an equation*                                          |
|                                                                                                                                          |
| -   **raq_18**: responses (1-5) to the question *R always crashes when I try to use it*                                                  |
|                                                                                                                                          |
| -   **raq_19**: responses (1-5) to the question *Everybody looks at me when I use R*                                                     |
|                                                                                                                                          |
| -   **raq_20**: responses (1-5) to the question *I can't sleep for thoughts of eigenvectors*                                             |
|                                                                                                                                          |
| -   **raq_21**: responses (1-5) to the question *I wake up under my duvet thinking that I am trapped under a normal distribution*        |
|                                                                                                                                          |
| -   **raq_22**: responses (1-5) to the question *My friends are better at R than I am*                                                   |
|                                                                                                                                          |
| -   **raq_23**: responses (1-5) to the question *If I am good at statistics people will think I am a nerd*                               |
+==========================================================================================================================================+
|                                                                                                                                          |
+------------------------------------------------------------------------------------------------------------------------------------------+

## Load packages and data

```{r}
remotes::install_github("dr-JT/semoutput")
remotes::install_github("profandyfield/discovr")
library(remotes)
library(corrplot) # nice correlation plots
library(easystats) # paramaters
library(tidyverse) # tidying visualization
library(psych) # fa func
```

```{r}
data <- discovr::raq
```

## Exploratory Factor Analysis

### Explore

1.  Explore the data.

> Any missing data?

```{r}

```

::: callout-note
## ANSWER
:::

> Outliers?

```{r}

```

::: callout-note
## ANSWER
:::

> Is there multicollinearity in the data?

```{r}

```

::: callout-note
## ANSWER
:::

> Plot a nice looking correlation matrix

```{r}

```

### Factor suitability

2.  Run and interpret the two diagnostic tests discussed in class to determine if factor analysis is appropriate as an analysis

```{r}

```

### Number of factors

3.  Determine how many factors to extract. Run `fa.parallel` (make sure you run a factor analysis (FA) and not PCA)

```{r}

```

> Looking at the graph. How many factors should we extract using:
>
> -   Eigenvalues \> 1?
>
> -   Scree plot?
>
> -   Parallel plot?

::: callout-note
## ANSWER
:::

> Now use the consensus method approach from `easystats`

```{r}

```

> Do they all agree?

::: callout-note
## ANSWER
:::

> How many factors should we extract? Justify your response.

::: callout-note
## ANSWER
:::

### Factor extraction

4.  Conduct a factor analysis using `psych::fa`(make sure to choose an extraction method and rotation method).

```{r}
#fa function

```

5.  Output the factors loadings in a nice table

```{r}

```

> Which items load onto which factor? Are there any items that seem to load strongly on multiple factors (i.e., items that are cross-loaded)?

::: callout-note
## ANSWER
:::

> Use `model_paramaters` and the `threshold` argument to get rid of loadings greater than .32.

```{r}

```

> Does this help with interpretation?

::: callout-note
## ANSWER
:::

> if not, try setting threshold to "max" in `model_paramters`

```{r}

```

> Better?

::: callout-note
## ANSWER
:::

### Naming

6.  Come up with names for your factors

::: callout-note
## ANSWER
:::

## Confirmatory factor analysis

7.  Split your data into a testing and training datasets

    ```{r}
    ```

8.  Fit the model you created on the test data using the the `cfa` function. How does the model fit fit? Use `semoutput (https://dr-jt.github.io/semoutput/)`

    ```{r}
    #look at fit indices discussed in class 
    ```

9.  Fit a competing model (come up with an alternative, nested, model)

    ```{r}

    ```

    > How does this model fit?

::: callout-note
## ANSWER
:::

10. Compare both models you fit above. Which one fits the data better?

    ```{r}
    #LRT 
    ```

::: callout-note
## ANSWER
:::

11. Write up the results. Be sure to include a figure of the final EFA and a table containing factor loadings, communality, uniqueness, and complexity
